Pollution, agricultural productivity, and development: Evidence from coal plants in India

### 2023-04-07 presentation/slides/slides.Rmd
changes to draft, analyses, and cleaning

---
title: "Pollution, agricultural productivity, and development: Evidence from coal in plants in India"
subtitle: 
author: 
  - "Joshua D. Merfeld <br> KDI School and IZA"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)

library(sf)
library(sp)
library(tidyverse)
library(gganimate)
library(transformr)
library(gifski)
library(ggthemes)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(lubridate)
library(icons)
library(png)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

@@

# Tables
plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yieldtable <- readRDS("pollution_development/draft/tables/yieldtable.rds")
yieldtablepm <- readRDS("pollution_development/draft/tables/yieldtablepm.rds")
yieldtabletwo <- readRDS("pollution_development/draft/tables/yieldtabletwo.rds")
yieldtablehet <- readRDS("pollution_development/draft/tables/yieldtablehet.rds")
pmeffects <- readRDS("pollution_development/draft/tables/yieldtablePM.rds")
labortable <- readRDS("pollution_development/draft/tables/labortable.rds")
labortablemonsoon <- readRDS("pollution_development/draft/tables/labortablemonsoon.rds")
labortablewinter <- readRDS("pollution_development/draft/tables/labortablewinter.rds")
labortableold <- readRDS("pollution_development/draft/tables/labortableold.rds")
ntltable <- readRDS("pollution_development/draft/tables/ntltable.rds")
yieldtableleads <- readRDS("pollution_development/draft/tables/yieldtableleads.rds")
yieldtablepostplant <- readRDS("pollution_development/draft/tables/yieldtablepostplant.rds")
laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")

@@

![]("../../../pollution_development/draft/tables/energy.png")

@@

![]("../../../pollution_development/draft/tables/energy.png")

@@

- Effects of pollution from coal plants on agricultural productivity in India
  - Previous evidence from gold mines (Aragon and Rud 2016) and workers in California (Graff Zivin and Neidell 2012)
- Higher exposure to pollution leads to lower agricultural productivity
  - Exposure defined as wind direction from coal plants towards villages
  - One additional day of exposure per season: decrease of 0.3 percent (yields)

@@

- Adults decrease labor allocation
  - Driven by large decrease in older half of sample; younger half increases allocation to farm labor
- More pollution in previous year also leads to lower nightlight growth
  - 25th to 75th percentile of exposure leads to decrease in nightlights of 0.7 percent

@@

@@

# Table
colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
kable(
      plantresultstable,
      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
      ) %>%
  footnote(
           "Note: Robust standard errors are in parentheses. The outcome in the first column is whether the village is within 30km of a coal plant in 1991. The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991. The last two columns are similarly defined, except using 2001 and 2011 as the years. * p<0.10 ** p<0.05 *** p<0.01",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE,
           escape = FALSE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(9, hline_after = TRUE) %>%
  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
  kable_classic_2()

@@

## Validating wind direction - particulate matter and exposure

# Table
colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pollutiontable,
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(3), bold = TRUE) %>%
  row_spec(6, hline_after = TRUE) %>%
  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
  kable_classic_2()

@@

## Effects of wind direction on agricultural productivity

# Table
colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtable,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(7, bold = TRUE) %>%
  row_spec(11, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

@@

# Table
colnames(yieldtablepm) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtablepm[-c(3:6),],
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  # footnote(
  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
  #          general_title = "",
  #          threeparttable = TRUE,
  #          footnote_as_chunk = TRUE
  #          ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(3, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

## Effects of PM on agricultural productivity, IV estimates

@@

# Table
colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtabletwo[-c(3:6, 12, 16:19),],
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  # footnote(
  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
  #          general_title = "",
  #          threeparttable = TRUE,
  #          footnote_as_chunk = TRUE
  #          ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(c(3,8), bold = TRUE) %>%
  row_spec(c(2), hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

@@

## Wind direction and labor allocation, by age

```{r labortableold, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
labortableold <- labortableold[,-1]
colnames(labortableold) <- c("all", "self", "wage", "farm", "non-farm")
kable(
      labortableold,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:6),width = "1.5cm") %>%
  row_spec(c(10), hline_after = TRUE) %>%
  row_spec(c(6,9), bold = TRUE) %>%
  kable_classic_2()

@@

## Large negative effects on older half

- Older half of sample decreases overall labor allocation
  - Driven by self-employment and nonfarm employment
- Younger half actually increases allocation!
- Consistent with health effects only on older adults, though not conclusive

@@

## Effects on aggregate growth?

- We know pollution affects productivity and labor allocation
- Might it also affect growth?
- Proxy for growth with nightlights
  - Look at previous year's exposure and current year's nightlights
  - "Growth" of nightlights
  - Village FE

@@

## Wind direction and growth of nightlights

@@

# Table
colnames(ntltable) <- c("(1)", "(2)")
kable(
      ntltable,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(c(9), hline_after = TRUE) %>%
  row_spec(c(5,8), bold = TRUE) %>%
  kable_classic_2()

@@

- Paper presents evidence that villages downwind from coal plants suffer in terms of agricultural productivity
  - Identification relies on within-village changes across seasons
- Wind exposure fluctuations lead to changes in agricultural productivity of around 2.4 percent based on mean absolute deviation
  - Instrumenting for particulate matter, effects are even larger (more than 20 percent)
- Clear changes in labor allocation
  - Oldest half of sample
- Slower growth of nightlights

@@

>>>>

---
title: "Pollution, agricultural productivity, and development: Evidence from coal plants in India"
subtitle: 
author: 
  - "Joshua D. Merfeld <br> KDI School and IZA"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
                      fig.width = 9, fig.height = 3.5, fig.retina = 3,
                      out.width = "100%",
                      cache = FALSE,
                      echo = TRUE,
                      message = FALSE, 
                      warning = FALSE,
                      hiline = TRUE
                      )

library(sf)
library(sp)
library(tidyverse)
library(gganimate)
library(transformr)
library(gifski)
library(ggthemes)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(lubridate)
library(png)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

@@

# Tables
plantresultstable <- readRDS("pollution_development/draft/tables/plantresultstable.rds")
pollutiontable <- readRDS("pollution_development/draft/tables/pollutiontable.rds")
yield1reducedform <- readRDS("pollution_development/draft/tables/yield1reducedform.rds")
yield2naive <- readRDS("pollution_development/draft/tables/yield2naive.rds")
yield3ivmain <- readRDS("pollution_development/draft/tables/yield3ivmain.rds")
yield4monthly <- readRDS("pollution_development/draft/tables/yield4monthly.rds")
yield5heterogeneity <- readRDS("pollution_development/draft/tables/yield5heterogeneity.rds")
yield6leads <- readRDS("pollution_development/draft/tables/yield6leads.rds")
yield7ivdiffindiff <- readRDS("pollution_development/draft/tables/yield7ivdiffindiff.rds")
labor1all <- readRDS("pollution_development/draft/tables/labor1all.rds")
labor2rural <- readRDS("pollution_development/draft/tables/labor2rural.rds")
labor3month <- readRDS("pollution_development/draft/tables/labor3month.rds")
laborplot <- readRDS("pollution_development/draft/tables/laborplot.rds")
monthplot <- readRDS("pollution_development/draft/tables/monthplot.rds")
opplot <- readRDS("pollution_development/draft/tables/opplot.rds")

@@

- india villages complain coal "power plant" (Adaniwatch)

@@

```{r energypng, echo = FALSE, message = FALSE, warning = FALSE}

knitr::include_graphics(
                        path = "energy.png",
                        auto_pdf = TRUE
                        )

@@

- Effects of pollution from coal plants on agricultural productivity in India
  - Previous evidence from gold mines (Aragon and Rud 2016) and workers in California (Graff Zivin and Neidell 2012)
- Higher exposure to pollution leads to lower agricultural productivity
  - Exposure defined as wind direction from coal plants towards villages
  - Also instrument for particulate matter 2.5 (PM 2.5)

@@

- Effects on labor allocation decisions, as well
- Largest effects on labor and yield in same month
  - Around very beginning of the monsoon season (June)

@@

## Empirical strategy

First stage

$$pm_{it} = \alpha^1_{i} + \gamma^1_{t} + \phi^1 wind_{it} + \psi^1 weather_{it} + \varepsilon^1_{it}$$

Second stage

$$yield_{it} = \alpha^2_{i} + \gamma^2_{t} + \phi^2 \hat{pm}_{it} + \psi^2 weather_{it} + \varepsilon^2_{it}$$

- i indexes villages (districts for labor)
- t indexes time (where time differs based on the analysis)
- standard errors clustered at village (or district) throughout
- common use of stars (one indicates 0.10) throughout

@@

# Table
colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
kable(
      plantresultstable,
      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(9, hline_after = TRUE) %>%
  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
  kable_classic_2()

@@

## Reduced form - wind and yield

# Table
colnames(yield1reducedform) <- c("(1)", "(2)", "(3)")
kable(
      yield1reducedform,
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2.5cm") %>%
  row_spec(13, bold = TRUE) %>%
  row_spec(15, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

## First stage - particulate matter and exposure, monthly level

# Table
colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pollutiontable,
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(3), bold = TRUE) %>%
  row_spec(6, hline_after = TRUE) %>%
  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
  kable_classic_2()

@@

## PM and agricultural productivity

# Table
colnames(yield2naive) <- c("(1)", "(2)", "(3)")
kable(
      yield2naive,
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2.5cm") %>%
  row_spec(5, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

@@

## Effects of PM on agricultural productivity (IV)

@@

# Table
colnames(yield3ivmain) <- c("(1)", "(2)", "(3)")
rownames(yield3ivmain)[8] <- "F (first stage)"
kable(
      yield3ivmain,
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2cm") %>%
  row_spec(5, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  #add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

## Heterogeneity

```{r heterogeneity, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield5heterogeneity) <- c(">p(50)", "<=p(50)", ">p(50)", "<=p(50)", "")
rownames(yield5heterogeneity)[c(1, 3)] <- c("PM 2.5", "PM times rain")
kable(
      yield5heterogeneity,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "4.5cm") %>%
  column_spec(c(2:6),width = "1.8cm") %>%
  row_spec(c(8), hline_after = TRUE) %>%
  row_spec(c(6), bold = TRUE) %>%
  add_header_above(c(" " = 1, "wind" = 2, "yield" = 2, " " = 1)) %>%
  kable_classic_2()

@@

## Wind direction and labor allocation, by rural status

```{r labortablerural, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labor1all) <- c("all", "all", "self", "wage", "farm", "non-farm")
colnames(labor2rural) <- c("all", "all", "self", "wage", "farm", "non-farm")

mat1 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat1) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat1) <- c("panel a: all")
mat2 <- matrix(c("", "", "", "", "", ""), nrow = 1)
colnames(mat2) <- c("all", "all", "self", "wage", "farm", "non-farm")
rownames(mat2) <- c("panel b: rural")

labortablerural <- rbind(mat1, labor1all[c(1:2, 7:8),], mat2, labor2rural[-c(3:6),], labor2rural[c(3:6),])

kable(
      labortablerural,
      align = "cccccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:7),width = "1.5cm") %>%
  row_spec(c(5,10), hline_after = TRUE) %>%
  row_spec(c(1,6,12), bold = TRUE) %>%
  kable_classic_2()

@@

## One last check

- Farm labor most common in July 
- Have monthly values of wind and pollution...
  - Heterogeneity by month of the season?
  - Monsoon season only

@@

## Heterogeneity by month, labor

```{r labortablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(labor3month) <- c("farm", "non-farm")
labor3month <- labor3month[-c(1:2, 13:14),]
labor3month <- labor3month[c(1:14),]
kable(
      labor3month,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(c(14), hline_after = TRUE) %>%
  row_spec(c(12), bold = TRUE) %>%
  kable_classic_2()

@@

## Heterogeneity by month, yield

```{r yieldtablemonthly, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# Table
colnames(yield4monthly) <- c("(1)", "(2)")
yield4monthly <- yield4monthly[c(1:14),]
kable(
      yield4monthly,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(12, bold = TRUE) %>%
  row_spec(14, hline_after = TRUE) %>%
  kable_classic_2()

## Coal plants or overall pollution?

- I use the location of coal plants to identify high-pollution areas
- However, some coal plants open part-way through sample
- Differences-in-differences combined with IV
  - Differences in coal pollution, specifically, and other forms of pollution

## Coal plants or overall pollution?

@@

# Table
colnames(yield7ivdiffindiff) <- c("(1)", "(2)", "(3)")
kable(
      yield7ivdiffindiff, caption = "Agricultural productivity and pollution source",
      align = "ccc", booktabs = TRUE, linesep = ""
      ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:4),width = "2cm") %>%
  row_spec(c(7), bold = TRUE) %>%
  row_spec(9, hline_after = TRUE) %>%
  kable_classic_2()

@@

- Paper presents evidence that villages downwind from coal plants suffer in terms of agricultural productivity
  - Identification relies on within-village changes across seasons
- Wind exposure fluctuations lead to changes in agricultural productivity of around 2.4 percent based on mean absolute deviation
  - Instrumenting for particulate matter, effects are even larger (more than 20 percent)
- Clear changes in labor allocation at beginning of the season
  - Yield only changes w.r.t. early-season pollution

@@

class: center, middle

<font size = "40"> Thank you! </font>

[https://joshmerfeld.github.io](https://joshmerfeld.github.io) <br>
Twitter: [@Josh\_Merfeld](twitter.com/Josh_Merfeld)
